The Symmetry of Partner Modelling

### 2016-02-04 rebuttal.tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript. As a follow-up of your comments, the manuscript underwent
several significant changes, that we summarize hereafter.

\begin{itemize}

    \item The title and the abstract have been completely changed to clarify the
        claims of the paper, which, as the editor pointed out, were not clear
        for some reviewers. In a nutshell, we don’t know what’s going on in he
        brain when building a shared solution, we leave this scientific debate
        to psycholinguists, but we observe how they vary (or not) across
        situations (across the 5 experiments)  and this enables some inferences.

    \item   Section 1, Introduction. We did minor edits except for the last
        paragraph has been completely rewritten it  that was somehow misleading
        regarding to the goals of the paper.  I think the point is now made
        clearly that the issue is not whether it is alway necessary to build a
        shared understanding, but that CSCL folds design tasks specifically for
        requiring shared understanding.

    \item We clarified the notion of grounding criterion.

    \item Section 2 has been split into partner modeling and mutual modeling (now
        section 3), to emphasized the fact that the question is the mutuality. The
        main research question is then clearly stated to address some of the
        reviewers cirtciisms

    \item Section 4. We clarified the status of the proposed formalism, which some
        reviewers find convenient and others criticize. Clearly, we are not
        proposing here a computational theory of mutual modeling. This notation is
        only useful because the text become quickly unreadable as one talks about
        the model that A build about the model that B build about A, while
        M(A,B,M(B,A,X)) is clear. So, it’s just a language for discussing
        experiments and hypotheses in a simple way. It’s function is internal to
        this paper

    \item For the 5 studies, we have re-written all the ‘discussion’ sections in a
        way that they cumulate naturally in alignment with the research questions

    \item As requested by the reviewers, we have completely rewritten  the
        conclusions to make them more explicit.  We cannot of course not pretend
        that our results are clearcut. We have to be careful not to publish only
        papers with significant results (statisticians have proved that some
        meta-analyses are biased by the fact that papers with significant results
        get more easily published). However, despite the fact that none of the 3
        hypothesis is declared as winner, the key results is that we prove that
        individual skills only explain a part of the variance, that the quality of
        social interactions matter significantly. This is why Gerry asked me (a
        few years ago I must say) to present these results that I mentionned in a
        ICLS panel.

\end{itemize}

\end{document}

### 2016-02-04 rebuttal.tex

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript and providing in-depth review of our submission.

The main points that were raised relate to:
\begin{itemize}
    \item the need for a better presentation of the concepts that we build upon,
    \item the need for a ``deeper synthetic discussion'', also ``accounting for the
        conflicting results'', that would present the ``theoretical
        picture'' resulting from our work,
    \item discuss a possible research agenda
\end{itemize}

In line with these comments and suggestions, the manuscript underwent several
significant changes that we summarize hereafter.

@@

\begin{itemize}

    \item The title and the abstract have been changed to clarify the
        claims of the paper, which, as the editor pointed out, were not clear
        for some reviewers. In a nutshell, we do not know what is going on in the
        brain when building a shared solution (we leave this scientific debate
        to psycholinguists), but we observe how they vary (or not) across
        situations (across the 5 experiments) and this enables some reflexions
        that we put in the context of collaborative learning.

    \item \emph{Section 1 -- Introduction} We did minor edits except for the last
        paragraph has been completely rewritten it  that was somehow misleading
        regarding to the goals of the paper.  We think the point is now made
        clearly that the issue is not whether it is alway necessary to build a
        shared understanding, but that CSCL folds design tasks specifically for
        requiring shared understanding.

    \item We clarified the notion of grounding criterion, as requested by
        reviewer 2 (page [TODO!!])

    \item \emph{Section 2} has been split into partner modeling and mutual modeling (now
        section 3) to emphasize the fact that the question is the mutuality. The
        main research question is then clearly stated to address some [WHICH?] of the
        reviewers' criticisms.

    \item \emph{Section 'Notations'} We clarified the status of the proposed formalism, which some
        reviewers found useful and others criticized. We are not
        proposing here a computational theory of mutual modeling. This notation is
        only useful because the text become quickly unreadable as one talks about
        the model that A build about the model that B build about A, while
        M(A,B,M(B,A,X)) is less ambiguous. We mainly see it as a language for discussing
        experiments and hypotheses in a simple way. Its function is internal to
        this paper.

    \item \emph{Section 'Studies'} For all the 5 studies, we have re-written all
        the \emph{discussion} sub-sections in a
        way that they naturally align with the research questions, leading to a
        clearer overall message.

    \item As requested by the reviewers, we have entirely rewritten  the
        conclusions to make them more explicit.  We cannot of course not pretend
        that our results are clearcut. We have to be careful not to publish only
        papers with significant results (statisticians have proved that some
        meta-analyses are biased by the fact that papers with significant results
        get more easily published). However, despite the fact that none of the 3
        hypothesis is declared as winner, the key results is that we prove that
        individual skills only explain a part of the variance, that the quality of
        social interactions matter significantly.

### 2016-02-05 mutual-modeling.tex
Polish section 4, 5, 6, 7.1

@@

\usepackage{xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\cf}{{\textit{cf\xspace}}}
\newcommand{\eg}{{\textit{e.g.\xspace}}}
\newcommand{\al}{{\textit{et al.\xspace}}}

@@

Natural language becomes cumbersome when describing things such as ``the model
that \A builds about the model that \B builds about A''. Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean \A has an
explicit, monolithic representation of \B: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, \A can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; it's usefulness is internal to this paper.  Besides, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item Tasks vary a lot with respect to how much they require mutual
        understanding.  The \emph{grounding criterion} -- denoted
        \groundingcriterion -- refers to how accurate the models need to be to
        succeed the task $T$. It can be calculated as the probability to succeed
        $T$ despite the fact $X$ is not grounded.
        $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
        between \Model{A}{B}{X} and the task performance.

    \item Before any specific grounding action, there is generally a non-null
        probability that $X$ is mutually understood by \A and \B (\eg $X$ is
        part of \A's and \B's cultures, it is manifest to co-present subjects
        or simply there is not much space for misunderstanding or disagreement
        about $X$). We simply could not collaborate without a certain level of
        initial grounds. We note the theoretical accuracy of \emph{initial
        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.

@@

Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$
about content $X$ at time $t_1$ during task T in order to increase
\Model{A}{B}{X} at time $t_2$ is the ratio between how much it is needed
((2)-(1)) and how much it costs (3)~\citep{traum1996miscommunication}:

\begin{multline} \label{eq:probrepair}
    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \simeq
    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
\end{multline}

This formula is presented as a qualitative summary, not as a real equation,
since several parameters are hard to quantify (\eg the cost of a communication
act depends upon the user as well). We use it for clarifying the parameters of
the experiments we present hereafter.

@@

        \item \model{A}{B}{actions} is about how well \A guesses what action \B has
            performed (study 2) or will perform next (study 1),

        \item \model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
            (study 3),

        \item \model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
            with respect to the material they learn together (study 4 and 5).

    \item $\mathcal{M}^{\circ}_{min}(A,B,X,T)$: our studies build upon various
        collaborative tasks: argumentation (study 3), games (study 1 and 2) and
        concept mapping (study 4 and 5). By varying the tasks, we do actually
        vary the grounding criterion but they all require a ``reasonably high''
        grounding criterion.  All tasks are about building a solution or a
        representation together. As stated earlier, we did not address everyday conversations.

    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university (and they hence had some general common
        grounds).  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a {\sc jigsaw} script.

@@

@@

Since a mental model is not directly observable,  the study of mutual modelling
is methodologically difficult. How can we for instance measure \Model{A}{B}{X}.
We proceed in two steps, first to capture \model{A}{B}{X} and then to estimate
\Model{A}{B}{X}.

{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask \A what he/she
believes about what \B knows, feels, intends to do, etc.  This raises obvious
methodological concerns since such a question triggers a modelling process
beyond what would naturally occur. To avoid this bias, one can estimate mutual
modelling after task completion.  Then, the obvious drawback are memory losses
and by post-hoc reconstruction.  The first method was used in study 2 and the
second one in the other studies. Another option would be to rely on external
behavioural metrics like eye-tracking: we hypothesise for instance that the
fixation time reflects the efforts engaged by the human to understand, hence,
model, the others. Such an approach has however not been investigated in the
presented studies.

{\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
Since we can only indirectly access it via what \B reports (\ie
\model{B}{B}{X}), accuracy can be estimated in 2 ways:

@@

    \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
        measuring if \A describes \B's emotions in the same way \B reports
        its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).

@@

Our method for investigating mutual modelling is observe the variations of
accuracy that result from variations of external parameters that appear as
variable in the formula presented in the previous section: the accuracy should
for go down if we increase the cost of grounding acts and goes up if we increase
the grounding criterion (the necessity to build a shared understanding). One way
to produce these variations is what CSCW researchers have called ``awareness
tools'', \ie functionalities that inform \A is about \B's actions that \A could
not directly perceive because \B was working on a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological tricks  to manipulate experimentally the  mutual modelling
activity.

\section*{Hypotheses and questions}

@@

As stated earlier, the fundamental challenge is to determine if mutual modelling
is an individual skill (hypothesis 1) or  the emergent property of social
interactions (hypothesis 3).  This question is empirically translated into the
symmetry of mutual modelling: what is the relationship between \Model{A}{B}{X}
and \Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that
mutual modelling is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$).
A high correlation might support $\mathcal{H}_{3}$ since there is a low
probability that randomly formed pairs integrate peers with the same level of
mutual modelling skills. However, a high correlation could also have another
explanation, stated in  $\mathcal{H}_{2}$: it could be that \A is good at
modelling \B and good at helping \B to model herself or himself.

@@

With triads, we may compute the accuracy of 6 models: \Model{A}{B}{X},
\Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X}, \Model{C}{B}{X} and
\Model{B}{C}{X}.This leads to two \emph{triangle questions}
(Fig.~\ref{mm_triangles}): Do \A and \B have the same accuracy when modelling
\C ($\Delta_2 = \Delta(\mathcal{M}^{\circ}(A,C,X),
\mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support ,
$\mathcal{H}_{2}$  (C has helped both \A and \B to model C) or support
$\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to
model each other accurately)

Conversely, does \C model more accurately \A than \B ($\Delta_3=
\Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
answer would support $\mathcal{H}_{1}$, \C being simply good at modelling any
partner.  It could also support  $\mathcal{H}_{3}$,  since the quality of
interactions influences the accuracy of two models that \C is building. A
negative answer would support $\mathcal{H}_{2}$ since $\Delta_3$ could mainly be
explained by the fact that  \A helped more \C to model him than \B did.

In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
whether the accuracy of mutual modelling depends more upon the modeller's effort
$\mathcal{H}_{1}$ or the modellee's behaviour  $\mathcal{H}_{2}$.

Please note that we consider the quality of interactions at triad level (ABC),
neglecting the cases where AB interact better for instance than BC, since there
was no ``private'' communication channel in the following studies. We do
nonetheless acknowledge that this point could be debated.

@@

We can go further by comparing self- versus other modelling ($\Delta_4$ in
Fig.~\ref{mm_rectangle}). A large difference would indicate that meta-cognitive
skills (self-modelling) and partner modelling skills are rather different
skills, while a small $\Delta_4$ could be interpreted as the indications that
these are two specific instance of a more general cognitive process. This
questions is however not central to this paper, and the value of $\Delta_4$ us
only available in one of the reported studies.

@@

We report five studies conducted  by different researchers in different contexts
between 2000 and 2015. They do not form a consistent research strategy but
actually the fact that some trends emerge despite their diversity constitutes
indeed the richness of this line of work.

\begin{table*}[ht!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

@@

{\sc SpaceMiners} is a 3D game that involves two players collecting minerals
located in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch
drones through the space after choosing their initial direction and speed. Once
launched, the trajectory of drones is only influenced by the gravity of planets
and by trajectory modification tools.  During the experiment, the teams were
confronted with three increasingly complex situations. The experiment was 2
hours long: a 30 minute tutorial and 3 levels of 30 minutes. Players uses using
a regular joystick and communicated with each other through an audio channel.

The independent variable was the availability of an awareness tool that shows to
player \A  the location and gaze direction of player \B. In the awareness
condition, players could switch to the \emph{scout mode} where they could view
what their partner was looking at. We hypothesize that this would enable
subjects to more accurately infer their teammate's intentions. Each player sat
in front of a distinct computer located in different rooms.

@@

In order to evaluate \gModel{A}{B} during the task, we used two questionnaires
(Fig.~\ref{study1:questionnaires}) that were displayed during each of the three
games, as a transparent layer appearing over the game display. The first
questionnaire concerned the player's intended actions. The second questionnaire
asked each player about what he thought her or his partner was intending to do.
Some answers were identical in both questionnaires (like ``adjusting a shot'')
while others were reversed (``to guide my partner'' versus ``to guide me'').
This method provides us with a subjective measure of accuracy
($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$) rather than an objective measure (\ie the
model \gmodel{A}{B} is compared to \B's next action) because some of the
activities proposed by the questionnaire were not observable by the environment
(\eg establishing a strategy). We calculated \Model{A}{B}{activity})
as the number of common answers between questionnaires \gmodel{A}{A} and
\gmodel{A}{B} in each game and computed the average value a across the 3 levels.

@@

\paragraph{Study-specific questions} The awareness tool permitted higher group
performance, but it did not improve the accuracy of the mutual model. Since
teams were free to use the awareness tool or not (the \emph{scout} mode), we
performed a post-hoc split of players depending on how much time they used it.
The split point was the mean of time spent in the scout mode and it led to the
constitution of two groups made up of 12 individuals ``short time in scout
mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
variance conducted on these contrasted groups revealed that pairs in the
awareness condition who spent more time in the scout mode reached higher levels
of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
does not support a causal direction. An alternative explanation could be that
good modellers are more social and hence appreciate the awareness tool.

@@

$\mathcal{H}_{3}$

@@

\paragraph{Grounding criterion} The correlation between \gModel{A}{B} and the
task performance (path lengths) was low: 0.15. Using a \emph{post-hoc} split on
\gModel{A}{B}, there is no significant difference between the performance of the
groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
\emph{post-hoc} split of the groups according to their performance did not show
any significant differences on \gModel{A}{B} ($F = 1.16, p = 0.29$).

@@

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modelling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge. 
Third, our results emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.

>>>>

\usepackage{booktabs}
\usepackage{array}

@@

\usepackage{xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\cf}{{\textit{cf}\xspace}}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\al}{{\textit{et al.}\xspace}}

@@

Natural language becomes cumbersome when describing things such as ``the model
that \A builds about the model that \B builds about A''. Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean \A has an
explicit, monolithic representation of \B: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, \A can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; its usefulness is internal to this paper.  Besides, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item Tasks vary a lot with respect to how much they require mutual
        understanding.  The \emph{grounding criterion} -- denoted
        \groundingcriterion -- represents the minimum level of modelling
        accuracy required for a task $T$ to succeed. Qualitatively, if the
        performance on a given task $T$ is correlated to \Model{A}{B}{X}, then
        \model{A}{B}{X} is significant to $T$ success, and the
        grounding criterion of $X$ for $T$
        ($\mathcal{M}^{\circ}_{min}(A,B,X,T)$) is non-zero.

        Under the assumption that the higher the correlation, the more
        critical \model{A}{B}{X} is to $T$, we hereafter use the correlation
        coefficient as an estimate of \groundingcriterion.
        \vspace{1em}

    \item Before any specific grounding action, there is generally a non-null
        probability that $X$ is mutually understood by \A and \B (\eg $X$ is
        part of \A's and \B's cultures, it is manifest to co-present subjects
        or simply there is not much space for misunderstanding or disagreement
        about $X$). We simply could not collaborate without a certain level of
        initial grounds. We note the theoretical accuracy of \emph{initial
        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
        \vspace{1em}

@@

Based on these 3 parameters, the probability of making an action $\alpha_{t}$
about content $X$ at time $t$ during task T in order to increase \Model{A}{B}{X}
is the ratio between how much it is needed and how much it
costs~\citep{traum1996miscommunication}:

\begin{equation}\label{eq:probrepair}
    p(\alpha_{t}(X,T) /_{\mathcal{M}^{\circ}_{t+1}(A,B,X)\nearrow}) \simeq
    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) -
    \mathcal{M}^{\circ}_{t}(A,B,X)}{cost (\alpha_{t})}
\end{equation}

This formula is presented as a qualitative summary, not as a real equation,
since several parameters are hard to quantify (\eg the cost of a communication
act depends upon the user as well).

We can further clarify the parameters in the context of the experiments we
present hereafter:

@@

        \item \Model{A}{B}{actions} is about how well \A guesses what action \B has
            performed (study 2) or will perform next (study 1),

        \item \Model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
            (study 3),

        \item \Model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
            with respect to the material they learn together (study 4 and 5).

    \end{enumerate}
        \vspace{1em}

    \item $\mathcal{M}^{\circ}_{min}(A,B,X,T)$: our studies build upon various
        collaborative tasks: argumentation (study 3), games (study 1 and 2) and
        concept mapping (study 4 and 5). By varying the tasks, we do actually
        vary the grounding criterion. The tasks were all designed to require a reasonably high
        grounding criterion, as they are meant for the participants to have to
        actually build a solution or a representation together.
        \vspace{1em}

    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university (and they hence had some general common
        grounds).  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a {\sc jigsaw} script.
        \vspace{1em}

@@

        While we introduce here formally the cost of grounding $cost(\alpha)$ as
        one relevant variable for the discussion of mutual modelling situations,
        we will not attempt to characterize it beyond these qualitative
        observations in the studies we present hereafter.

@@

Since a mental model is not directly observable,  the study of mutual modelling
is methodologically difficult. How can we for instance measure \Model{A}{B}{X}?
We proceed in two steps, first to capture \model{A}{B}{X} and then to estimate
\Model{A}{B}{X}.

{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask \A what he/she
believes about what \B knows, feels, intends to do, etc.  This raises obvious
methodological concerns since such a question triggers a modelling process
beyond what would naturally occur. To avoid this bias, one can estimate mutual
modelling after task completion. Then, the obvious drawback are memory losses
and post-hoc reconstruction. The first method was used in study 2 and the
second one in the other studies. Another option would be to rely on external
behavioural metrics like eye-tracking: we hypothesise for instance that the
fixation time reflects the efforts engaged by the human to understand, hence,
model, the others. Such an approach has however not been investigated in the
presented studies.

{\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
access the reference model \refmodel{B}{X} to estimate its accuracy.
Since we can only indirectly access it via what \B reports (\ie
\model{B}{B}{X}), accuracy can be estimated in 2 ways:

@@

    \item Subjective accuracy: In study 3, for instance, we compute \Model{A}{B}{X} by
        measuring if \A describes \B's emotions in the same way \B reports
        its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).

@@

Our method for investigating mutual modelling relies on the observation of the variations of
accuracy that result from variations of external parameters (the variables of
the formula~\ref{eq:probrepair} above): for instance, the accuracy should
go down if we increase the cost of grounding acts, and conversly go up if we increase
the grounding criterion (\ie the necessity to build a shared understanding). One way
to produce these variations relies what CSCW researchers call ``awareness
tools'', \ie functionalities that inform \A about \B's actions that \A can
not directly perceive due to \B working in a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological levers to experimentally manipulate the mutual modelling
activity.

\section{Hypotheses and questions}

@@

As stated earlier, the fundamental challenge is to determine if mutual modelling
is an individual skill (hypothesis $\mathcal{H}_{1}$ below) or  the emergent
property of social interactions (hypothesis $\mathcal{H}_{3}$).  This question
is empirically translated into the symmetry of mutual modelling
(Fig.~\ref{mm_symmetry}): what is the relationship between \Model{A}{B}{X} and
\Model{B}{A}{X}? A low symmetry would mean that mutual modelling is mainly an
individual attitude/aptitude ($\mathcal{H}_{1}$).  A high correlation might
support $\mathcal{H}_{3}$ since there is a low probability that randomly formed
pairs integrate peers with the same level of mutual modelling skills. However, a
high correlation could also have another explanation, stated in
$\mathcal{H}_{2}$: it could be that \A is good at modelling \B and good at
helping \B to model herself or himself.

@@

With triads, we may compute the accuracy of 6 models: \Model{A}{B}{X},
\Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X}, \Model{C}{B}{X} and
\Model{B}{C}{X}. This leads to two \emph{triangle questions}
(Fig.~\ref{mm_triangles}): Do \A and \B have the same accuracy when modelling
\C ($\Delta_2 = \Delta(\mathcal{M}^{\circ}(A,C,X),
\mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support ,
$\mathcal{H}_{2}$  (C has helped both \A and \B to model C) or support
$\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to
model each other accurately)

Conversely, does \C model with similar accuracy \A and \B? (low $\Delta_3=
\Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
answer would support $\mathcal{H}_{1}$, \C being simply good at modelling any
partner.  It could also support  $\mathcal{H}_{3}$,  since the quality of
interactions influences the accuracy of two models that \C is building. A
negative answer would support $\mathcal{H}_{2}$ since $\Delta_3$ could mainly be
explained by the fact that  \A helped more \C to model him than \B did.

In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
whether the accuracy of mutual modelling depends more upon the modeller's effort
($\mathcal{H}_{1}$) or the modellee's behaviour ($\mathcal{H}_{2}$).

Note that we consider the quality of interactions at triad level (\A, \B, \C),
neglecting the cases where \A and \B interact better for instance than \B and \C, since there
was no ``private'' communication channel in the following studies. We do
nonetheless acknowledge that this point could be debated.

@@

We can go further by comparing self- versus other modelling ($\Delta_4$ in
Fig.~\ref{mm_rectangle}). A large difference would indicate that meta-cognitive
skills (self-modelling) and partner modelling skills are rather different
skills, while a small $\Delta_4$ could be interpreted as the indications that
these are two specific instance of a more general cognitive process. This
questions is however not central to this paper, and the value of $\Delta_4$ is
only available in one of the reported studies.

@@

We report on five studies conducted  by different researchers in different contexts
between 2000 and 2015. They do not form a consistent research strategy but
actually the fact that some trends emerge despite their diversity constitutes
indeed the richness of this line of work.

\begin{table*}[ht!]
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\centering
%\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.1cm}L{1.65cm}L{1.7cm}L{1.8cm}L{1.6cm}L{1.6cm}}
\toprule
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ 
\midrule
\textbf{Task}                 & game in virtual space            & game in physical space          & building an argument map  & building a concept map    & building a concept map       \\
\midrule
\textbf{Interactions}         & audio                            & written                         & audio / video             & audio                     & audio                        \\
\midrule
\textbf{Shared editor}        & 3D space                         & 2D map                          & concept map               & concept map               & concept map                  \\
\midrule
\textbf{Group size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\midrule
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\midrule
\textbf{Awareness tool}       & partner's position               & partner's current/past pos.     & -                         & partner's concept map     & partner's scores at pre-test \\
\midrule
\textbf{Dependent variable}       & team/indiv. performance            & team/indiv. performance           & team performance          & team/indiv. knowledge       & team/indiv. knowledge          \\
\midrule
\textbf{Independent variable}     & awareness tool \textit{vs} none  & awareness tool \textit{vs} none & audio \textit{vs} audio+video & script \textit{vs} none & awareness tool \textit{vs} none       \\
%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\bottomrule
\end{tabular}
%}
\caption{Overview of the studies}
\label{synthesis_table}
\end{table*}

@@

{\sc SpaceMiners} is a 3D game that involves two players harvesting ore
found on asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch
drones through the space after choosing their initial direction and speed. Once
launched, the trajectory of drones is influenced by the gravity of planets
and by ``trajectory modification'' tools. During the experiment, the teams were
confronted with three increasingly complex situations. The experiment was 2
hours long: a 30 minute tutorial and 3 levels of 30 minutes. The players were using
a regular joystick and communicated with each other through an audio channel.

The independent variable was the availability of an awareness tool that shows to
player \A  the location and gaze direction of player \B: in this ``awareness''
condition, players could switch to the \emph{scout mode} where they could view
what their partner was looking at. We hypothesize that this would enable
subjects to more accurately infer their teammate's intentions. Each player sat
in front of a distinct computer located in different rooms.

@@

In order to evaluate \gModel{A}{B} during the task, we used two questionnaires
(Fig.~\ref{study1:questionnaires}) that were displayed during each of the three
games, as a transparent layer appearing over the game display. The first
questionnaire concerned the player's intended actions. The second questionnaire
asked each player about what he thought her or his partner was intending to do.
Some answers were identical in both questionnaires (like ``adjusting a shot'')
while others were reversed (``to guide my partner'' versus ``to guide me'').
This method provides us with a subjective measure of accuracy
($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$) rather than an objective measure (\ie the
model \gmodel{A}{B} is compared to \B's next action) because some of the
activities proposed by the questionnaire were not observable by the environment
(\eg establishing a strategy). We calculated \Model{A}{B}{activity})
as the number of common answers between questionnaires \gmodel{A}{A} and
\gmodel{A}{B} in each game and computed the average value across the 3 levels.

@@

\paragraph{Study-specific questions} The awareness tool permitted higher group
performance, but it did not improve the accuracy of the mutual model. Since
teams were free to use the awareness tool or not (the \emph{scout} mode), we
performed a post-hoc split of players depending on how much time they used it.
The split point was the mean of time spent in the scout mode and it led to the
constitution of two groups made up of 12 individuals ``short time in scout
mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
variance conducted on these contrasted groups revealed that pairs in the
awareness condition who spent more time in the scout mode reached higher levels
of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a post-hoc split
does not support a causal direction. An alternative explanation could be that
good modellers are more social and hence appreciate the awareness tool.

@@

@@

\paragraph{Grounding criterion} The correlation between \gModel{A}{B} and the
task performance (path lengths) was low: 0.15. Using a post-hoc split on
\gModel{A}{B}, there is no significant difference between the performance of the
groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
post-hoc split of the groups according to their performance did not show
any significant differences on \gModel{A}{B} ($F = 1.16, p = 0.29$).

@@

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modelling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge. 
Third, our results emerge from a post-hoc reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.
